Diverse Genotype-by-Weather Interactions in Switchgrass

### 2024-03-12 manuscript/revision2/revision2.tex
manuscript: AH edits from 2024-03-07

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Mapping Mixtures of Genotype-by-Weather Interactions in Switchgrass (Panicum virgatum)},
  pdfauthor={Alice H. MacQueen; Li Zhang; Samuel A. Smith; Jason Bonnette; Arvid R. Boe; Phillip A. Fay; Felix B. Fritschi; David B. Lowry; Robert B. Mitchell; Francis M. Rouquette Jr; Yanqi Wu; Arbel Harpak; Thomas E. Juenger},
  pdfkeywords={allele-by-environment effect variation, antagonistic
pleiotropy, photoperiod, cumulative rainfall, genetic variation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\title{Mapping Mixtures of Genotype-by-Weather Interactions in
Switchgrass (Panicum virgatum)}

@@

\begin{abstract}
The timing of vegetative and reproductive growth in plants
(``phenological timings'') depends on both genetic variation and
environmental cues; thus, phenological traits can have important
genotype-by-environment (GxE) interactions. Identifying the mixture and
prevalence of GxE primed by weather prior to the phenological event
(GxWeather) should aid prediction and manipulation of phenological
timings. Here, we map GxWeather effects on phenological timings in two
highly divergent switchgrass (\emph{Panicum virgatum}) populations using
repeated plantings of cloned individuals from these populations at eight
sites spanning the central United States. We distinguish GxWeather
patterns covarying with interpretable weather-based cues from agnostic,
site-based patterns. Most GxWeather effects belong to the latter
category and do not covary with weather-based cues. However, 65\% of
effects on the timing of vegetative growth in the Gulf population covary
with daylength 14 days prior to green-up date. 33\% of effects on the
timing of flowering in the Gulf population covary with cumulative
rainfall in the seven days prior to flowering, while 22\% of effects on
flowering timing in the Midwest population covary with day length change
two days prior to flowering. An independent pseudo-F2 cross of Gulf and
Midwest individuals at the same sites mapped eight additive QTLs for
flowering and three for vegetative growth timing, all with significant
associations and three with enrichment of highly significant
associations in our model of GxWeather effects. We demonstrate that we
can identify genetic variation with GxWeather and assign these loci to
specific weather-based cues or other patterns. Breeding for particular
alleles at these loci could change flowering responsiveness to
photoperiod and rainfall cues in switchgrass. More broadly, this
approach could be used to identify genetic marker-environment
interactions in any species with related populations phenotyped in
multiple environments.
\end{abstract}

@@

\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
\firstpage[8]{2}

Plant phenological timings are major components of plant fitness
affected by multiple external environmental cues (e.g.~degree of winter
chilling, day length, temperature, and water availability) that signal
existing or upcoming growing conditions (1--3). Genetic responses to
environmental cues determine the speed, timing, and energy apportioned
to vegetative and reproductive growth and shape both the individual's
lifespan and its lifetime production of viable seed. Day length (or
photoperiod) is one of the most predictable environmental cues, and
genetic sensitivity to photoperiod protects plants from potentially
fatal consequences of phenological responses to temperature cues at the
``wrong'' time of year. However, the usefulness of specific
environmental cues depends on both features of the environment, such as
cue predictability and relevance, and the species' adaptive strategies
(4). Species with wide natural distributions can have multiple distinct
environmentally cued phenological responses: for example, populations of
sunflower (\emph{Helianthus annuus}) exhibit day-neutral, facultative
short day, and facultative long-day flowering responses, which vary with
their environments (5, 6). Distinct genetic responses in different
environments are known as genotype by environment interactions, or GxE.

Flowering time, or the transition from vegetative to reproductive
growth, is a common subject of GxE research (5--11), a key output of
selection driving adaptation to local environments (2, 12, 13), and a
selection target for crop improvement to adapt crops to local or future
environments (14). Changing flowering responsiveness to photoperiod cues
has allowed geographic range expansion and increased yields in several
cereal species (15--19) and other crops (20, 21). Recent statistical
advances in studying phenological GxE have involved determining critical
environmental indices before the phenological event occurs, such as
photothermal time within a critical growth window (10). However, most
studies of flowering GxE focus on finding a single, best fitting form of
genotype-environment covariance, despite the key expectation that
different genetic subpopulations, and even different genomic regions,
have likely evolved distinct patterns of GxE. Additionally, despite
theoretical predictions that local adaptation should involve
antagonistic pleiotropy, or sign-changing GxE, at the level of
individual loci (22--25), previous work has found limited evidence of
antagonistic pleiotropy (12, 26). However, this work has been limited by
a known statistical bias that reduced detection of genetic effects that
differ in sign (26--28). Thus, despite substantial interest in the
frequencies of various forms of GxE, the prevalence of antagonistic
pleiotropy relative to other forms of GxE remains unknown.

Previous research suggests that switchgrass phenological timings should
have GxWeather and that these timings could differ by genetic
subpopulation. Switchgrass is considered a short-day plant with
reproductive development strongly linked to day of the year (29).
However, as part of its wide environmental adaptation across the eastern
half of North America, its photoperiodicity has been predicted to differ
by plant latitude of origin (30, 31). We previously found divergent
Midwest and Gulf genetic subpopulations of switchgrass with distinct
sets of environmental adaptations, in that both populations had distinct
genetic variation associated with each of two fitness proxies, biomass
and survival (32). The Midwest genetic subpopulation is primarily
composed of individuals from the well-studied upland switchgrass ecotype
(33, 34), while the Gulf subpopulation has individuals from the
well-studied lowland ecotype and the phenotypically intermediate coastal
ecotype (32).

Here, we estimate GxWeather for two phenological timings in switchgrass
by assigning patterns of genetic effects on phenology across gardens to
many patterns of weather covariance at these gardens. To do this, we
phenotype a diversity panel of hundreds of switchgrass genotypes from
the Midwest and Gulf subpopulations for the timing of vegetative and
reproductive development at eight common garden locations spanning 17
degrees of latitude. These gardens cover the majority of the latitudinal
and climatic range of switchgrass and capture the most comprehensive
picture to date of the environmental variation this species encounters.
We define multiple ways phenological timings might covary with weather
(@tbl-covar) and additional ways phenological timings might vary by site
(@SI-info), then jointly re-estimate genetic effects on these timings at
all eight sites using the set of these covariance matrices that
significantly improved the modeled log-likelihood when included
(@SI-info) (35). We use the Bayesian framework \emph{mash} (multivariate
adaptive shrinkage), developed by (35), to refine effect size estimates
from genome-wide association (GWAS) conducted on individuals at each of
the eight sites. \emph{mash} allows us to identify and specify multiple
covariance structures among genetic effect estimates across sites,
including structures that represent covariance in weather variables of
interest. Importantly, \emph{mash} does not have a statistical bias in
detecting genetic effects with the same or opposite signs (36). To
confirm our genetic mapping of GxWeather, we compare our posterior
genetic effect estimates from \emph{mash} to mapping results from an
outbred pseudo-F2 cross grown at the same sites. Our analyses allow us
to describe the weather cues and genetic variation affecting phenology
in two divergent natural populations of switchgrass.

@@

Narrow-sense heritabilities (h\textsuperscript{2}) indicated that
rank-changing GxE for these phenotypes was present across the common
gardens (Figure~\ref{fig-map} C). h\textsuperscript{2} were typically
high at individual gardens: 59\% on average for green-up date, and 87\%
for flowering date. However, h\textsuperscript{2} were variable across
gardens, and green-up dates were uncorrelated (r\textsuperscript{2}
\textless{} 0.2) or negatively correlated between pairs of gardens
(Figure~\ref{fig-map} B). These negative and small correlations
undoubtedly contributed to the low h\textsuperscript{2} values for
green-up and flowering date when estimated jointly at all eight gardens:
h\textsuperscript{2} was 0.8\% for green-up and 23.2\% for flowering
date.

@@

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects across all eight common gardens using the mash model,
which allows effects to covary between gardens in many ways
(Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). We specified three
categories of covariance structures across gardens: ``canonical''
covariance, with simple patterns of effect size covariance introduced in
the initial \emph{mash} manuscript; ``data-driven'' covariances derived
from common patterns of SNP effects observed in the data, and
``GxWeather'' covariance, estimated from the covariance of empirical
weather patterns at each garden at specific times before the
phenological event (Table 1, SI Appendix, Section S2). For example, the
more similar that the amount of rainfall prior to flowering is for
genotypes A and B between gardens in MI and MO, the nearer these
covariances will be to one. If rainfalls are uncorrelated, covariance
will be near zero; if there is rank changing (e.g.~high rainfall before
flowering of A at MO but not MI, and high rainfall before flowering for
B at MI but not MO), this covariance will be near negative one. We
specified multiple covariance structures for the canonical, data-driven,
and GxWeather categories, each of which represented a specific pattern
of covariance in genetic effect size estimates that could be included in
the model. The phenotypic correlations for the timing of vegetative
growth had moderate negative correlations between the Texas and North
gardens, particularly in the Gulf subpopulation and Both subpopulations
(Figure~\ref{fig-map} B). If these phenotypic correlations have a
genetic component, they could be partially or completely captured by the
covariance structures specified in the mash model.

@@

As the climate and the natural environment change, it is increasingly
critical to understand how patterns of gene-environment and
plant-environment interactions will change in response. To do this, we
must understand the current patterns of trait covariation across
environments, the genetic underpinnings of these patterns, and the cases
where this covariation can be altered. Here, we demonstrate that we can
associate multiple patterns of GxWeather with specific genomic regions
using a switchgrass diversity panel grown at eight common gardens. We
can assign genetic effects to both GxWeather patterns with interpretable
weather-based cues, and to agnostic, site-based patterns. We use this
approach to study GxWeather for the timings of vegetative and
reproductive development in the deeply genetically diverged Gulf and
Midwest subpopulations of switchgrass.

Our analysis of the timing of vegetative growth in the Gulf and in Both
subpopulations revealed substantial antagonistic pleiotropy in effects
between the Texas and North gardens (Figure~\ref{fig-effects} B). This
result supports theoretical models that local adaptation should involve
antagonistic pleiotropy at the level of individual loci (22--25), and is
the first experimental work using GWAS across common gardens to find
antagonistic pleiotropy to be common in small genomic regions (12, 37,
38).

@@

Identifying the environmental cues that are predictive of, or even
correlated with, plant phenotypic responses remains a major challenge to
studies interrogating gene action across many natural environments. The
GxWeather photoperiod and cumulative rainfall cues we identify here are
functions of the genotypes measured and capture only a minority of SNP
effects on flowering. We know still less about the overwintering
parameters that cause variation in the start of vegetative growth. We
could only assign SNP effects to a GxWeather covariance structures in
four of the six phenotype \& genetic subpopulations we modeled. More
generally, it is difficult to predict the time scales over which
individuals may integrate environmental cues, particularly in perennial
species which may integrate these cues over longer time scales. If this
integration time itself varies between individuals, we cannot select a
covariance structure that reflects this, though these structures would
likely be highly correlated with GxWeather structures we did include.
Mash offers an opportunity to specify multiple environmental cues and
compete them to explain patterns of genetic effects, allowing us to
detect how important these cues are genome-wide, and how strongly each
cue influences each SNP. This is a key development to further improve
our understanding of genetic variation in GxE.

@@

\caption{\label{fig-map}Characterization of green-up and flowering dates
from the switchgrass diversity panel. (a) Map and trait histograms of
green-up and flowering dates across two genetically distinct switchgrass
subpopulations and eight common gardens. Purple represents individuals
from the Midwest genetic subpopulation, and pink individuals from the
Gulf subpopulation. Vertical dashed lines indicate the summer solstice.
Common gardens are arranged in latitudinal order. (b) Phenotypic
correlations between clonal replicates planted at eight common gardens,
within and between two genetic subpopulations. (c) Narrow sense
heritability of green-up and flowering within single common gardens
(purple) and across all eight common gardens (green), within and between
two genetic subpopulations. (d) Flow diagram of the methods applied to
the green-up and flowering dates to jointly estimate SNP effects across
all sites. Mash was fit to SNP effect data and used to find covariance
matrices that improved the mash model likelihood using a large set of
randomly selected, relatively unlinked SNP effects; this model was
applied to a ``strong'' set of SNP effects with large effect sizes in
the univariate GWAS.}

@@

\includegraphics{images/Figure_2_Covariance_matrices_and_posterior_weights.png}

@@

The formation and resequencing of the diversity panel has been described
previously (32). The diversity panel contained 134 sequenced, clonally
propagated individuals from the Midwest genetic subpopulation, and 229
from the Gulf genetic subpopulation. To allow for the possibility that
different subpopulations had different strengths of connection between
our phenotypes and genotypes (38), we conducted three sets of genetic
analyses: on Gulf and Midwest genotypes separately, and on both
subpopulations together (`Both'). Analyses to determine narrow-sense
heritability (h\textsuperscript{2}) for green-up and flowering were done
using linear mixed models and followed (32). Details on these models can
be found in (SI Appendix, Section S3,S4).

@@

To evaluate the prevalence and kinds of covariance patterns of SNP
effects across our common gardens, we used multivariate adaptive
shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
(35). To do this, we first conducted univarate genome-wide association
(GWAS) on site-specific best linear unbiased predictors (BLUPs) for the
timing of vegetative growth and flowering, using the switchgrassGWAS
package and the methods in (39). We used the effect estimates for single
nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect
estimates for both traits at each garden. We then jointly modeled
genetic effect estimates, and the mixture of ways these effects might
covary, across all eight common gardens, using a two-part procedure. We
first modeled genetic effects with different combinations of covariance
structures, using the effect size estimates from a subset of \ul{19K}
relatively unlinked (r\textsuperscript{2} \textless{} 0.2), randomly
selected SNPs used in the GWAS. In these models (hereafter referred to
as ``random''), we then specified a variety of potential covariance
structures among genetic effect size estimates from each garden (as
expanded on below) and used a greedy algorithm to iteratively select and
add covariance structures that significantly improved the log likelihood
of the \emph{mash} model. Given the high correlation between some of the
covariance patterns we tested, this approach allowed us to select only
the subset of patterns of covariance that significantly improved the fit
of the jointly inferred structure of genetic effect estimates present in
the data. We then selected 19K relatively unlinked (r\textsuperscript{2}
\textless{} 0.2) SNPs from each univariate GWAS that had the largest
effect estimates in each GWAS (hereafter ``strong'' effects), which gave
sets of \ul{\textbf{48K - 80K}} SNPs, and jointly re-estimated these
effect size estimates at all eight sites using the set of significant
covariance matrices that most improved the model fit on the random
effects. Notably, the majority of these ``strong'' effects were not
significant in univariate GWAS.

@@

Using the lfsr rather than the local false discovery rate (lfdr) is a
critical change in our ability to detect antagonistic pleiotropy. The
lfdr, like other measures of FDR, focuses on if we have enough evidence
to reject the null hypothesis that an effect j is 0, or that there is a
significant effect. Previous studies of antagonistic pleiotropy
(e.g.~(37)) have used the lfdr or equivalent statistical tests to detect
antagonistic pleiotropy. These tests were conservative, in that they
required two non-zero effects of different signs, while tests for
differential sensitivity required only one non-zero effect. This
previous work recognized that this testing bias could lead to
undercounting occurrences of antagonistic pleiotropy (26, 27), and
sought to reduce it by permutation (28). However, using the lfsr to test
for antagonistic pleiotropy does not undercount occurrences of
antagonistic pleiotropy, as this statistic answers a fundamentally
different question. For each effect j, the lfsr\textsubscript{j} is
defined as the probability that we make an error in the sign of effect j
if we were forced to declare the effect positive or negative (36). Thus,
rather than asking ``Are these two effects different?'' - as we
reasonably expect two effects to be, even if this difference cannot be
measured - the local false sign rate answers a more meaningful question:
Can we be confident in the sign of this effect?

@@

To confirm candidate genomic regions and patterns of allelic effects
found in the diversity panel, we analyzed flowering in an outbred
pseudo-F2 cross between four individuals, two Midwest and two Gulf
individuals. The formation of this mapping population has been described
previously (34); additional details on QTL mapping can be found in SI
Appendix, Section S6. To be directly comparable to the diversity panel
data, only 2019 phenology data from the pseudo-F2 cross from the same
eight common garden sites were used.

@@

\bibitem[\citeproctext]{ref-levene1953genetic}
\CSLLeftMargin{22. }%
\CSLRightInline{H. Levene, Genetic equilibrium when more than one
ecological niche is available. \emph{The American Naturalist}
\textbf{87}, 331--333 (1953).}

\bibitem[\citeproctext]{ref-felsenstein1976theoretical}
\CSLLeftMargin{23. }%
\CSLRightInline{J. Felsenstein, The theoretical population genetics of
variable selection and migration. \emph{Annual review of genetics}
\textbf{10}, 253--280 (1976).}

\bibitem[\citeproctext]{ref-kawecki2004conceptual}
\CSLLeftMargin{24. }%
\CSLRightInline{T. J. Kawecki, D. Ebert, Conceptual issues in local
adaptation. \emph{Ecology letters} \textbf{7}, 1225--1241 (2004).}

\bibitem[\citeproctext]{ref-hedrick1986genetic}
\CSLLeftMargin{25. }%
\CSLRightInline{P. W. Hedrick, Genetic polymorphism in heterogeneous
environments: A decade later. \emph{Annual review of ecology and
systematics} \textbf{17}, 535--566 (1986).}

\bibitem[\citeproctext]{ref-des2013genotype}
\CSLLeftMargin{26. }%
\CSLRightInline{D. L. Des Marais, K. M. Hernandez, T. E. Juenger,
Genotype-by-environment interaction and plasticity: Exploring genomic
responses of plants to the abiotic environment. \emph{Annual Review of
Ecology, Evolution, and Systematics} \textbf{44}, 5--29 (2013).}

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Diverse Genotype-by-Weather Interactions in Switchgrass},
  pdfauthor={Alice H. MacQueen; Li Zhang; Samuel A. Smith; Jason Bonnette; Arvid R. Boe; Phillip A. Fay; Felix B. Fritschi; David B. Lowry; Robert B. Mitchell; Francis M. Rouquette Jr; Yanqi Wu; Arbel Harpak; Thomas E. Juenger},
  pdfkeywords={allele-by-environment effect variation, antagonistic
pleiotropy, photoperiod, cumulative rainfall, genetic variation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\title{Diverse Genotype-by-Weather Interactions in Switchgrass}

@@

\begin{abstract}
The timing of vegetative and reproductive growth in plants
(``phenological timings'') depends on genetic effects, environmental
(e.g., weather) cues, and possibly their interaction. Quantifying these
GxWeather interactions can aid prediction and manipulation of
phenological timings. Here, we map GxWeather effects on phenological
timings in two highly divergent switchgrass (\emph{Panicum virgatum})
populations using repeated plantings of cloned individuals from these
populations at eight sites across the central United States. Some
genetic effects covary with interpretable weather-based cues. For
example, in the Gulf population, 65\% of genetic effects on the timing
of vegetative growth covary with daylength 14 days prior to green-up
date, and 33\% of genetic effects on the timing of flowering covary with
cumulative rainfall in the seven days prior to flowering. However, most
variation in genetic effects across sites is site-specific and can not
be attributed to variation in weather variables alone. We demonstrate
that we can identify genetic variation with GxWeather and assign these
loci to specific weather-based cues or other patterns. Breeding for
particular alleles at the GxWeather loci we identified could change
flowering responsiveness in a photoperiod and rainfall-specific way.
More broadly, our approach proposes a refined, powerful characterization
of genotype-by-environment interactions in any species phenotyped in
multiple environments.
\end{abstract}

@@

\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
\firstpage[15]{2}

Plant phenological timings are major components of plant fitness
affected by multiple external environmental cues (e.g.~degree of winter
chilling, day length, temperature, and water availability) that signal
existing or upcoming growing conditions (1--3). Genetic responses to
environmental cues determine the speed, timing, and energy apportioned
to vegetative and reproductive growth and shape both the individual's
lifespan and its lifetime production of viable seed. Day length (or
photoperiod) is one of the most predictable environmental cues, and
genetic sensitivity to photoperiod protects plants from potentially
fatal consequences of phenological responses to temperature cues at the
``wrong'' time of year. However, the usefulness of specific
environmental cues depends on both features of the environment, such as
cue predictability and relevance, and the species' adaptive strategies
(4). Species with wide natural distributions can have multiple distinct
environmentally cued phenological responses: for example, populations of
sunflower (\emph{Helianthus annuus}) exhibit day-neutral, facultative
short day, and facultative long-day flowering responses, which vary with
their environments (5, 6). Distinct genetic responses in different
environments are known as genotype-by-environment interactions, or GxE.

Flowering time, or the transition from vegetative to reproductive
growth, is a common subject of GxE research (5--11), a key output of
selection driving adaptation to local environments (2, 12, 13), and a
selection target for crop improvement to adapt crops to local or future
environments (14). Changing flowering responsiveness to photoperiod cues
has allowed geographic range expansion and increased yields in several
cereal species (15--19) and other crops (20, 21). Recent statistical
advances in studying phenological GxE have involved determining critical
environmental indices before the phenological event occurs, such as
photothermal time within a critical growth window (10). However, most
studies of flowering GxE focus on finding a single, best fitting form of
genotype-environment covariance, despite the key expectation that
different genetic subpopulations, and even different genomic regions,
have likely evolved distinct patterns of GxE (22). Additionally, despite
theoretical predictions that local adaptation should involve
antagonistic pleiotropy, or sign-changing GxE, at the level of
individual loci (23--26), previous empirical work has found limited
evidence of antagonistic pleiotropy (12, 27). However, this work has
been limited by a known statistical bias that reduced detection of
genetic effects that differ in sign (27--29). Thus, despite substantial
interest in the frequencies of various forms of GxE, the prevalence of
antagonistic pleiotropy relative to other forms of GxE remains unknown.

Previous research suggests that switchgrass phenological timings should
have GxWeather and that these timings could differ by genetic
subpopulation. Switchgrass is considered a short-day plant with
reproductive development strongly linked to day of the year (30).
However, as part of its wide environmental adaptation across the eastern
half of North America, its photoperiodicity has been predicted to differ
by plant latitude of origin (31, 32). We previously found divergent
Midwest and Gulf genetic subpopulations of switchgrass with distinct
sets of environmental adaptations, in that both populations had distinct
genetic variation associated with each of two fitness proxies, biomass
and survival (33). The Midwest genetic subpopulation is primarily
composed of individuals from the well-studied upland switchgrass ecotype
(34, 35), while the Gulf subpopulation has individuals from the
well-studied lowland ecotype and the phenotypically intermediate coastal
ecotype (33).

Here, we estimate GxWeather for two phenological timings in switchgrass
by assigning patterns of genetic effects on phenology across gardens to
many patterns of weather covariance at these gardens. To do this, we
phenotype a diversity panel of hundreds of switchgrass genotypes from
the Midwest and Gulf subpopulations for the timing of vegetative and
reproductive development at eight common garden locations spanning 17
degrees of latitude. These gardens cover the majority of the latitudinal
and climatic range of switchgrass and capture the most comprehensive
picture to date of the environmental variation this species encounters.
We define multiple ways phenological timings might covary with weather
(@tbl-covar) and additional ways phenological timings might vary by site
(@SI-info), then jointly re-estimate genetic effects on these timings at
all eight sites using the set of these covariance matrices that
significantly improved the modeled log-likelihood when included
(@SI-info) (36). We use the Bayesian framework \emph{mash} (multivariate
adaptive shrinkage) developed by (36), to refine effect size estimates
from genome-wide association (GWAS) conducted on individuals at each of
the eight sites. \emph{mash} allows us to identify and specify multiple
covariance structures among genetic effect estimates across sites,
including structures that represent covariance in weather variables of
interest. Importantly, this method circumvents statistical biases in
detecting genetic effects with the same or opposite signs (37). To
confirm our genetic mapping of GxWeather, we compare our posterior
genetic effect estimates from \emph{mash} to mapping results from an
outbred pseudo-F2 cross grown at the same sites. Our analyses allow us
to describe the weather cues and genetic variation affecting phenology
in two divergent natural populations of switchgrass.

@@

Narrow-sense heritabilities (h\textsuperscript{2}) suggested that
rank-changing GxE for these phenotypes was present across the common
gardens (Figure~\ref{fig-map} C). h\textsuperscript{2} were typically
high at individual gardens: 59\% on average for green-up date, and 87\%
for flowering date. However, h\textsuperscript{2} were variable across
gardens, and green-up dates were uncorrelated (r\textsuperscript{2}
\textless{} 0.2) or negatively correlated between pairs of gardens
(Figure~\ref{fig-map} B). These negative and small correlations
undoubtedly contributed to the low h\textsuperscript{2} values for
green-up and flowering date when estimated jointly at all eight gardens:
h\textsuperscript{2} was 0.8\% for green-up and 23.2\% for flowering
date.

@@

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects across all eight common gardens using the mash model,
which can flexibly capture many modes of covariance across gardens
(Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). We specified three
qualitative categories of covariance structures across gardens:
``canonical'' covariance, with simple patterns of effect size covariance
introduced in the initial \emph{mash} manuscript; ``data-driven''
covariances derived from common patterns of SNP effects observed in the
data, and ``GxWeather'' covariance, estimated from the covariance of
empirical weather patterns at each garden at specific times before the
phenological event (Table 1, SI Appendix, Section S2). For example, the
more similar that the amount of rainfall prior to flowering is for
genotypes A and B between gardens in MI and MO, the nearer these
covariances will be to one. If rainfalls are uncorrelated, covariance
will be near zero; if there is rank changing (e.g.~high rainfall before
flowering of A at MO but not MI, and high rainfall before flowering for
B at MI but not MO), this covariance will be near negative one. We
specified multiple covariance structures for the canonical, data-driven,
and GxWeather categories, each of which represented a specific pattern
of covariance in genetic effect size estimates that could be included in
the model. The phenotypic correlations for the timing of vegetative
growth had moderate negative correlations between the Texas and North
gardens, particularly in the Gulf subpopulation and Both subpopulations
(Figure~\ref{fig-map} B). If these phenotypic correlations have a
genetic component, they could be partially or completely captured by the
covariance structures specified in the mash model.

@@

As the climate and the natural environment change, it is increasingly
critical to understand how patterns of plant-environment interactions
will change in response. To do this, we must understand the current
patterns of trait covariation across environments, the genetic
underpinnings of these patterns, and the cases where this covariation
can be altered. Here, we demonstrate that we can associate multiple
patterns of GxWeather with specific genomic regions using a switchgrass
diversity panel grown at eight common gardens. We can assign genetic
effects to both GxWeather patterns with interpretable weather-based
cues, and to agnostic, site-based patterns. We use this approach to
study GxWeather for the timings of vegetative and reproductive
development in the deeply genetically diverged Gulf and Midwest
subpopulations of switchgrass.

Our analysis of the timing of vegetative growth in the Gulf and in Both
subpopulations revealed substantial antagonistic pleiotropy in effects
between the Texas and North gardens (Figure~\ref{fig-effects} B). This
result supports theoretical models that local adaptation should involve
antagonistic pleiotropy at the level of individual loci (23--26), and is
the first experimental work using GWAS across common gardens to find
antagonistic pleiotropy to be common in small genomic regions (12, 38,
39).

@@

Identifying the environmental cues that are predictive of, or even
correlated with, plant phenotypic responses remains a major challenge to
studies interrogating gene action across many natural environments. The
GxWeather photoperiod and cumulative rainfall cues we identify here are
functions of the genotypes measured and capture only a minority of SNP
effects on flowering. We know still less about the overwintering
parameters that cause variation in the start of vegetative growth. We
could only assign SNP effects to a GxWeather covariance structures in
four of the six phenotype \& genetic subpopulations we modeled. More
generally, it is difficult to predict the time scales over which
individuals may integrate environmental cues, particularly in perennial
species which may integrate these cues over longer time scales. If this
integration time itself varies between individuals, we cannot select a
covariance structure that reflects this, though these structures would
likely be highly correlated with GxWeather structures we did include.
Our approach offers an opportunity to specify multiple environmental
cues and compete them to explain patterns of genetic effects, allowing
us to detect how important these cues are genome-wide, and how strongly
each cue influences each SNP. This is a key development to further
improve our understanding of genetic variation in GxE.

@@

\caption{\label{fig-map}Characterization of green-up and flowering dates
from the switchgrass diversity panel. (a) Map and trait histograms of
green-up and flowering dates across two genetically distinct switchgrass
subpopulations and eight common gardens. Purple represents individuals
from the Midwest genetic subpopulation, and pink individuals from the
Gulf subpopulation; map positions represent the original collection
locations for the genotypes, and shapes represent the ecotype of the
genotype. Vertical dashed lines indicate the summer solstice. Common
gardens are arranged in latitudinal order. (b) Phenotypic correlations
between clonal replicates planted at eight common gardens, within and
between two genetic subpopulations. (c) Narrow sense heritability of
green-up and flowering within single common gardens (purple) and across
all eight common gardens (green), within and between two genetic
subpopulations. (d) Flow diagram of the methods applied to the green-up
and flowering dates to jointly estimate SNP effects across all sites.
Mash was fit to SNP effect data and used to find covariance matrices
that improved the mash model likelihood using a large set of randomly
selected, relatively unlinked SNP effects; this model was applied to a
``strong'' set of SNP effects with large effect sizes in the univariate
GWAS.}

@@

\includegraphics{Figure_2_recolor_covariance_matrices_and_posterior_weights.png}

@@

The formation and resequencing of the diversity panel has been described
previously (33). The diversity panel contained 134 sequenced, clonally
propagated individuals from the Midwest genetic subpopulation, and 229
from the Gulf genetic subpopulation. To allow for the possibility that
different subpopulations had different strengths of connection between
our phenotypes and genotypes (38), we conducted three sets of genetic
analyses: on Gulf and Midwest genotypes separately, and on both
subpopulations together (`Both'). Analyses to determine narrow-sense
heritability (h\textsuperscript{2}) for green-up and flowering were done
using linear mixed models and followed (33). Details on these models can
be found in (SI Appendix, Section S3,S4).

@@

To evaluate the prevalence and kinds of covariance patterns of SNP
effects across our common gardens, we used multivariate adaptive
shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
(35). To do this, we first conducted univarate genome-wide association
(GWAS) on site-specific best linear unbiased predictors (BLUPs) for the
timing of vegetative growth and flowering, using the switchgrassGWAS
package and the methods in (40). We used the effect estimates for single
nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect
estimates for both traits at each garden. We then jointly modeled
genetic effect estimates, and the mixture of ways these effects might
covary, across all eight common gardens, using a two-part procedure. We
first modeled genetic effects with different combinations of covariance
structures, using the effect size estimates from a subset of \ul{19K}
relatively unlinked (r\textsuperscript{2} \textless{} 0.2), randomly
selected SNPs used in the GWAS. In these models (hereafter referred to
as ``random''), we then specified a variety of potential covariance
structures among genetic effect size estimates from each garden (as
expanded on below) and used a greedy algorithm to iteratively select and
add covariance structures that significantly improved the log likelihood
of the \emph{mash} model. Given the high correlation between some of the
covariance patterns we tested, this approach allowed us to select only
the subset of patterns of covariance that significantly improved the fit
of the jointly inferred structure of genetic effect estimates present in
the data. We then selected 19K relatively unlinked (r\textsuperscript{2}
\textless{} 0.2) SNPs from each univariate GWAS that had the largest
effect estimates in each GWAS (hereafter ``strong'' effects), which gave
sets of \ul{\textbf{48K - 80K}} SNPs, and jointly re-estimated these
effect size estimates at all eight sites using the set of significant
covariance matrices that most improved the model fit on the random
effects. Notably, the majority of these ``strong'' effects were not
significant in univariate GWAS.

@@

Using the lfsr rather than the local false discovery rate (lfdr) is a
critical change in our ability to detect antagonistic pleiotropy. The
lfdr, like other measures of FDR, focuses on if we have enough evidence
to reject the null hypothesis that an effect j is 0, or that there is a
significant effect. Previous studies of antagonistic pleiotropy
(e.g.~(37)) have used the lfdr or equivalent statistical tests to detect
antagonistic pleiotropy. These tests were conservative, in that they
required two non-zero effects of different signs, while tests for
differential sensitivity required only one non-zero effect. This
previous work recognized that this testing bias could lead to
undercounting occurrences of antagonistic pleiotropy (26, 27), and
sought to reduce it by permutation (28). However, using the lfsr to test
for antagonistic pleiotropy does not undercount occurrences of
antagonistic pleiotropy, as this statistic answers a fundamentally
different question. For each effect j, the lfsr\textsubscript{j} is
defined as the probability that we make an error in the sign of effect j
if we were forced to declare the effect positive or negative (37). Thus,
rather than asking ``Are these two effects different?'' - as we
reasonably expect two effects to be, even if this difference cannot be
measured - the local false sign rate answers a more meaningful question:
Can we be confident in the sign of this effect?

@@

To confirm candidate genomic regions and patterns of allelic effects
found in the diversity panel, we analyzed flowering in an outbred
pseudo-F2 cross between four individuals, two Midwest and two Gulf
individuals. The formation of this mapping population has been described
previously (35); additional details on QTL mapping can be found in SI
Appendix, Section S6. To be directly comparable to the diversity panel
data, only 2019 phenology data from the pseudo-F2 cross from the same
eight common garden sites were used.

@@

\bibitem[\citeproctext]{ref-Weine2023.06.21.545998}
\CSLLeftMargin{22. }%
\CSLRightInline{E. Weine, S. P. Smith, R. K. Knowlton, A. Harpak,
Tradeoffs in modeling context dependency in complex trait genetics.
\emph{bioRxiv} (2024)
https:/doi.org/\href{https://doi.org/10.1101/2023.06.21.545998}{10.1101/2023.06.21.545998}.}

\bibitem[\citeproctext]{ref-levene1953genetic}
\CSLLeftMargin{23. }%
\CSLRightInline{H. Levene, Genetic equilibrium when more than one
ecological niche is available. \emph{The American Naturalist}
\textbf{87}, 331--333 (1953).}

\bibitem[\citeproctext]{ref-felsenstein1976theoretical}
\CSLLeftMargin{24. }%
\CSLRightInline{J. Felsenstein, The theoretical population genetics of
variable selection and migration. \emph{Annual review of genetics}
\textbf{10}, 253--280 (1976).}

\bibitem[\citeproctext]{ref-kawecki2004conceptual}
\CSLLeftMargin{25. }%
\CSLRightInline{T. J. Kawecki, D. Ebert, Conceptual issues in local
adaptation. \emph{Ecology letters} \textbf{7}, 1225--1241 (2004).}

\bibitem[\citeproctext]{ref-hedrick1986genetic}
\CSLLeftMargin{26. }%
\CSLRightInline{P. W. Hedrick, Genetic polymorphism in heterogeneous
environments: A decade later. \emph{Annual review of ecology and
systematics} \textbf{17}, 535--566 (1986).}
